---
pagetitle: "Schools in Trafford"
lang: "en-GB"
output: 
  html_document:
    css: styles.css
    highlight: null
    mathjax: null
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(crosstalk) ; library(tidyverse) ; library(sf) ; library(leaflet) ; library(leaflet.extras) ; library(reactable) ; library(shiny) ; library(htmlwidgets) ; library(htmltools)

# local authority boundary
sf <- st_read("data/local_authority.geojson", quiet = TRUE)

# schools
df <- read_csv("data/schools.csv") %>% 
  mutate(phase = fct_relevel(as_factor(phase),
                             level = c("Primary", "Secondary", "Further Education", "Continuous")),
         ofsted_rating = fct_relevel(as_factor(ofsted_rating),
                              level = c("Outstanding", "Good", "Requires improvement", "Serious Weaknesses")),
         website = ifelse(is.na(website), "", str_c("<a href='", website, "'target='_blank'>", str_remove(website, "http://www.|https://www.|www.|http://|https://"), "</a>")),
         popup = ifelse(is.na(ofsted_rating), str_c("<strong>", name, "</strong><br/><a href='", compare_performance, "'target='_blank'>Compare performance with other schools</a>") %>% map(HTML), str_c("<strong>", name, "</strong><br/><a href='", ofsted_report, "'target='_blank'>Latest Ofsted report</a><br/><a href='", compare_performance, "'target='_blank'>Compare performance with other schools</a>") %>% map(HTML)))  %>% 
  arrange(name)

# shared dataframe
sd <- SharedData$new(df, group = "name")
# subset of shared dataframe
sd_table <- select(df, name, headteacher, website, nursery_provision, sixth_form) %>%
  SharedData$new(group = "name")
```

```{r, map}
map <- leaflet(sd, width = "100%", height = 350) %>%
  addTiles(urlTemplate = "https://{s}.basemaps.cartocdn.com/light_all/{z}/{x}/{y}{r}.png", attribution = '&copy; <a href="https://www.openstreetmap.org/copyright">OpenStreetMap</a> contributors &copy; <a href="https://carto.com/attributions">CARTO</a> | <a href="https://www.ons.gov.uk/methodology/geography/licences">Contains OS data © Crown copyright and database right (2020)</a> | Data: <a href="https://get-information-schools.service.gov.uk" target="_blank">DfE</a>', options = tileOptions(minZoom = 11, maxZoom = 18)) %>%
  addPolygons(data = sf, fillColor = "#CCCCCC", weight = 0.8, opacity = 1, color = "#212121") %>%  
  addAwesomeMarkers(popup = df$popup, icon = ~makeAwesomeIcon(icon = "fas fa-school", library = "fa", iconColor = "#FFFFFF", markerColor = "purple")) %>%
  addFullscreenControl()
```

```{r, table}
# improve accessibility
options(reactable.language = reactableLang(
  noData = "No schools found",
  pagePreviousLabel = "Previous page",
  pageNextLabel = "Next page"))

table <- reactable(
  sd_table,
  resizable = TRUE,
  borderless = TRUE,
  wrap = FALSE,
  searchable = TRUE,
  highlight = TRUE,
  selection = "multiple",
  onClick = "select",
  rowStyle = list(cursor = "pointer"),
  columns = list(
    name = colDef(name = "School",
                  minWidth = 190,
                  style = list(position = "sticky", left = 0, background = "#fff", zIndex = 1),
                  headerStyle = list(position = "sticky", left = 0, background = "#fff", zIndex = 1)),
    headteacher = colDef(name = "Headteacher",
                         minWidth = 150),
    website = colDef(name = "Website", 
                     minWidth = 170,
                     html = TRUE),
    nursery_provision = colDef(name = "Nursery", 
                               cell = JS("function(cellInfo) {return cellInfo.value === 'No' ? '\u2718' : '\u2713'}")),
    sixth_form = colDef(name = "Sixth form",
                        cell = JS("function(cellInfo) {return cellInfo.value === 'No' ? '\u2718' : '\u2713'}")))
)
```
<header>
```{r, logo}
a(
  href = "https://www.trafforddatalab.io", 
  img(src = "https://www.trafforddatalab.io/assets/logo/trafforddatalab_logo.svg", 
      alt = "Trafford Data Lab", 
      width = "100",
      style = "position:absolute; top:0; right:0; padding: 10px;")
  )
```
</header><main>
```{r, ui}
div(class = "container-fluid",
    h1("Schools in Trafford", style = "padding-top: 25px; margin-bottom: 20px;"),
    div(class = "row",
        div(class = "col-xs-12 col-sm-6 col-md-6 col-lg-6", map),
        div(class = "col-xs-6 col-sm-3 col-md-3 col-lg-3", 
            list(
              # Phase
              filter_checkbox(id = "phase", label = "Phase", sharedData = sd, group = ~phase),
              # Type
              filter_checkbox(id = "type", label = "Type", sharedData = sd, group = ~type))),
        div(class = "col-xs-6 col-sm-3 col-md-3 col-lg-3", 
            list(
              # Gender
              filter_checkbox(id = "gender", label = "Gender", sharedData = sd, group = ~gender, inline = TRUE),
              br(),
              # Pupil count
              filter_slider(id = "number_of_pupils", label = "Number of pupils", sharedData = sd, column = ~number_of_pupils, step = 10, ticks = FALSE),
              br(),
              # FSM
              filter_slider(id = "percent_fsm", label = "Free School Meals", sharedData = sd, column = ~percent_fsm, step = 1, round = TRUE, sep = "", ticks = FALSE, post = "%"),
              br(),
              # Ofsted rating
              filter_checkbox(id = "ofsted_rating", label = "Ofsted rating", sharedData = sd, group = ~ofsted_rating)
              ))),
    br(),
    h2("School information"),
    div(class = "row",
        div(class = "col-sm-12", table)),
    div(class = "row",
        div(class = "col-sm-12", 
            p("Data were retrieved on 11 June 2020 from the ", a("Department for Education's", href = 'https://www.gov.uk/government/organisations/department-for-education'), a("register of schools and colleges.", href = 'https://get-information-schools.service.gov.uk'), "The number of pupils and the percentage of those on free school meals are drawn from a school audit conducted in January 2019."), 
            p("For more information about schools in Trafford, please visit ", a("trafford.gov.uk/residents/schools", href = 'https://www.trafford.gov.uk/residents/schools', target = "_blank"))),
        br()
        )
    )
```
</main>
<script>
  // Remove the crosstalk selection button from the Leaflet map. Map may not have initialised at this point so set up an interval to keep calling until it has
  var removeSelectionButtonFn = setInterval(removeSelectionButton, 100);
  
  function removeSelectionButton() {
    try {
      // We don't have an id or name so have to do it by classname and then check we have the correct button by testing the 'title' attribute
      var arrEasyButtonContainer = document.getElementsByClassName('easy-button-container');
  
      for (var i = 0; i < arrEasyButtonContainer.length; i++) {
        if (arrEasyButtonContainer[i].childNodes[0].title == "Make a selection") {
          var parent = arrEasyButtonContainer[i].parentNode;
          parent.removeChild(arrEasyButtonContainer[i]);    // remove the whole container plus button
          clearInterval(removeSelectionButtonFn);  // cancel any further calls to this function as we have completed all actions
        }
      }
    }
    catch(e) {
      clearInterval(removeSelectionButtonFn);   // something went wrong so cancel fn call
    }
  }
  
  // Add aria-label attribute to the fullscreen Leaflet map button to identify its purpose as the link is empty (assessibility issue)
  // Map may not have initialised at this point so set up an interval to keep calling until it has
  var updateFullscreenButtonFn = setInterval(updateFullscreenButton, 100);
  
  function updateFullscreenButton() {
    try {
      // We don't have an id or name so have to do it by classname
      var arrFullscreenButtonLink = document.getElementsByClassName('leaflet-control-fullscreen-button');
      
      if (arrFullscreenButtonLink.length > 0) {
        // set aria-label to describe the function of the link/button
        arrFullscreenButtonLink[0].setAttribute('aria-label', 'Maximise/minimise map.');
        
        clearInterval(updateFullscreenButtonFn);    // cancel any further calls to this function as we have completed all actions
      }
    }
    catch(e) {
      clearInterval(updateFullscreenButtonFn);    // something went wrong so cancel fn call
    }
  }
  
  // Add type="hidden" attribute to all form elements with classname "irs-hidden-input" to remove errors in
  // WAVE due to the inputs not having labels.
  // The controls may not have initialised at this point so set up an interval to keep calling until they have
  var addInputHiddenFn = setInterval(addInputHidden, 500);
  
  function addInputHidden() {
    try {
      var arrHiddenInputs = document.getElementsByClassName('irs-hidden-input');
    
      if (arrHiddenInputs.length > 0) {
        for (var i = 0; i < arrHiddenInputs.length; i++) {
          arrHiddenInputs[i].setAttribute('type', 'hidden');
        }
        
        clearInterval(addInputHiddenFn);   // cancel any further calls to this function as we have completed all actions
      }
    }
    catch(e) {
      clearInterval(addInputHiddenFn);   // something went wrong so cancel fn call
    }
  }
  
  // Remove label tags created by crosstalk for the checkboxes and sliders as they cause validation errors,
  // and replace them with spans containing the text of the label
  // Reason for the error is that the 'for' attribute isn't referencing an id of a form control
  // NOTE: the do while loop always references array index 0 because as the elements are removed from the DOM they are also removed from the array
  var arrFormLabels = document.getElementsByClassName('control-label');
  var parent;
  var labelText;
  var newNode;
  
  do {
    parent = arrFormLabels[0].parentNode;       // container of the label
    labelText = arrFormLabels[0].firstChild.textContent;    // get the label text we want to preserve
    
    newNode = document.createElement('span');   // create the replacement element to act as the heading...
    newNode.appendChild(document.createTextNode(labelText));   // ...with the former label text
    newNode.setAttribute('class', 'formGroupHeading');    // add CSS class to make it look like the label
    
    parent.replaceChild(newNode, arrFormLabels[0]);   // replace the old label with the new span
  }
  while (arrFormLabels.length > 0);
  
  // Add aria-labels to slider controls to inform the user as to their purpose
  var arrSliders = document.getElementsByClassName('crosstalk-input-slider');
  
  for (var i = 0; i < arrSliders.length; i++) {
    arrSliders[i].setAttribute('aria-label', 'Slider range control');
  }
</script>